---
title: Dow's Trick Or Treat
author: Richard Ian Carpenter
date: '2018-10-31'
slug: dow-s-trick-or-treat
categories:
  - dow jones
  - market
tags:
  - dow jones
  - market
description: ''
topics: []
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Loading packages:
library(lubridate)
library(blogdown)
library(kableExtra)
library(tidyverse)
library(gridExtra)
library(knitr)
library(moments)

# Setting working directory and reading in the data:
setwd("~/R/blog_new")
df <- read.csv("~/R/oct2018_dji_data.csv")

# Lubridate and summary stats:
df$date <- ymd(df$date)
# summary(df)

# Creating some fields...
# ... first is the daily difference:
df$daily_diff <- c(0, diff(df$close))
# ... a column of the first day's closing amount:
df$first_close <- df$close[1]
# ... this is the level change, with the first day set at zero:
df$level_change <- df$close - df$first_close
# ... this is the percent change, with the first day set at zero:
df$pct_change <- round((df$close/df$first_close),2)

# Simple plots and another summary statistic:
p1 <- ggplot(data = df, aes(x = date, y = close)) + geom_line() +
    theme_bw() +
    labs(title = "Dow Jones Index, October 2018",
         subtitle = "Market closing price",
         x = "Date", y = "Closing Price", 
         caption = "Source: Yahoo! Finance")
p2 <- ggplot(data = df, aes(x = date, y = daily_diff)) + geom_line() +
    theme_bw() +
    geom_hline(aes(yintercept = 0), color = "blue") +
    labs(title = "Dow Jones Index, October 2018",
         subtitle = "Daily change in market closing price",
         x = "Date", y = "Difference in Closing Price", 
         caption = "Source: Yahoo! Finance")
p3 <- ggplot(data = df, aes(x = date, y = pct_change)) + geom_line() +
    theme_bw() +
    geom_hline(aes(yintercept = 1.0), color = "blue") +
    labs(title = "Dow Jones Index, October 2018",
         subtitle = "Percent price change; Indexed, 1 Oct 2018 = 1.00",
         x = "Date", y = "Percent", 
         caption = "Source: Yahoo! Finance")
p4 <- ggplot(data = df, aes(x = date, y = level_change)) + geom_line() +
    theme_bw() +
    geom_hline(aes(yintercept = 0), color = "blue") +
    labs(title = "Dow Jones Index, October 2018",
         subtitle = "Actual price change; Indexed, 1 Oct 2018 = 0",
         x = "Date", y = "Amount", 
         caption = "Source: Yahoo! Finance")

# Histogram:
p5 <- ggplot(data = df, aes(x = daily_diff)) + geom_histogram(bins = 50)
```

The Dow Jones Industrial Average (DJIA) gave us an interesting, if not a little scary, performance this month.  When we look at the difference in closing prices, we can see that it was mostly within a range of +/-500 points:

```{r, echo = FALSE}
grid.arrange(p1, p2, nrow = 2)
```

Another view is the percent change in the DJIA, indexing on 1 October 2018 and seeing how it performed:

```{r, echo = FALSE}
p3
```

``` {r, echo = FALSE}
df2 <- read.csv("~/R/ytd2018_dji_data.csv")

# Lubridate and summary stats:
df2$date <- ymd(df2$date)
# summary(df2)

# Creating some fields...
# ... first is the daily difference:
df2$daily_diff <- c(0, diff(df2$close))
# ... a column of the first day's closing amount:
df2$first_close <- df2$close[1]
# ... this is the level change, with the first day set at zero:
df2$level_change <- df2$close - df2$first_close
# ... this is the percent change, with the first day set at zero:
df2$pct_change <- round((df2$close/df2$first_close),2)

# Simple plots and another summary statistic:
p6 <- ggplot(data = df2, aes(x = date, y = close)) + geom_line() +
    theme_bw() +
    geom_hline(aes(yintercept = 25000), color = "blue") +
    labs(title = "Dow Jones Index, 2 Jan 2018 to 31 Oct 2018",
         subtitle = "Market closing price",
         x = "Date", y = "Closing Price", 
         caption = "Source: Yahoo! Finance")
p7 <- ggplot(data = df2, aes(x = date, y = daily_diff)) + geom_line() +
    theme_bw() +
    geom_hline(aes(yintercept = 0), color = "blue") +
    labs(title = "Dow Jones Index, 2 Jan 2018 to 31 Oct 2018",
         subtitle = "Daily change in market closing price",
         x = "Date", y = "Difference in Closing Price", 
         caption = "Source: Yahoo! Finance")
p8 <- ggplot(data = df2, aes(x = date, y = pct_change)) + geom_line() +
    theme_bw() +
    geom_hline(aes(yintercept = 1.0), color = "blue") +
    labs(title = "Dow Jones Index, 2 Jan 2018 to 31 Oct 2018",
         subtitle = "Percent price change; Indexed, 2 Jan 2018 = 1.00",
         x = "Date", y = "Percent", 
         caption = "Source: Yahoo! Finance")
p9 <- ggplot(data = df2, aes(x = date, y = level_change)) + geom_line() +
    theme_bw() +
    geom_hline(aes(yintercept = 0), color = "blue") +
    labs(title = "Dow Jones Index, 2 Jan 2018 to 31 Oct 2018",
         subtitle = "Actual price change; Indexed, 2 Jan 2018 = 0",
         x = "Date", y = "Amount", 
         caption = "Source: Yahoo! Finance")

# Histogram:
p10 <- ggplot(data = df2, aes(x = daily_diff)) + geom_histogram(bins = 50)
```

On a similar note, I also looked at the year-to-date performance of the DJIA and noticed that volatility in the closing price occured in two periods: the one we are experiencing now in October and the other that occured in the February to April time period.  The May - September time period were relatively quiet in terms of the DJIA closing price:

```{r, echo = FALSE}
grid.arrange(p6, p7, nrow = 2)
```

Likewise, the performance appeared to be range bound of +/-5% when indexed to the first trading day of the year (2 January 2018):

```{r, echo = FALSE}
p8
```

Something to build on from here would be looking at the components of the DJIA and compare their individual performances over the same time period.  I am working on a Shiny app since it is more dynamic than this blog.  Once it is complete, I will include a link to it.

This is all I have for the moment.  I hope you enjoyed the (brief) post. 