---
title: Continued Introduction
author: Richard Ian Carpenter
date: '2018-06-19'
slug: continued-introduction
categories:
  - intro
tags:
  - intro
description: ''
topics: []
---

This is a continuation to my first post just to see how things look using `blogdown`.

So... first, I'll call one of the built in data sets:

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(hexSticker)

data(cars)
summary(cars)
```

Now that the summary statistics are out of the way, here's a quick plot using the `ggplot2` package, part of the `tidyverse`:

```{r, echo = FALSE}
p1 <- ggplot(data = cars, aes(x = speed, y = dist)) +
    geom_point() +
    geom_smooth(method = 'lm', se = FALSE)

p1
```

Finally, here's my site logo.  I created this using the `hexSticker` package:

```{r, echo = FALSE}
s1 <- sticker(p1, package = "jadg", p_size=18, s_x=1, s_y=1, s_width=1.2, s_height=1, p_x = 1, p_y = 1.7, h_color = "white", h_fill = "black", url = "www.justadatageek.com", u_size = 6, u_color = "white")

s1
```

That's it for tonight.  As I get more comfortable with how this all works, I'll start blogging.  Plus, I'll be migrating my previous blog, which gives me some instant content.

Stay tuned!
