---
title: The 2018 Dow Jones Industrial Average's Performance
author: Richard I. Carpenter
date: '2019-01-01'
slug: the-2018-dow-jones-industrial-average-s-performance
categories:
  - dow jones
  - market
tags:
  - dow jones
  - market
description: ''
topics: []
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Working Blog Script #3
# Actual blogdown script has not been created
# Date Created: 13 NOV 2018
# Date Updated: 01 JAN 2019

# Loading packages:
library(lubridate)
library(blogdown)
library(kableExtra)
library(tidyverse)
library(gridExtra)
library(knitr)
library(moments)

# Reading in the data:
df <- read.csv("~/R/dji_data_2018.csv")

# Lubridate and summary stats:
df$date <- ymd(df$date)
# summary(df)

# Creating some fields...
# ... first is the daily difference:
df$daily_diff <- c(0, diff(df$close))
# ... a column of the first day's closing amount:
df$first_close <- df$close[1]
# ... this is the level change, with the first day set at zero:
df$level_change <- df$close - df$first_close
# ... this is the percent change, with the first day set at zero:
df$pct_change <- round((df$close/df$first_close),2)
# ... this is to make the volume amounts more manageable with ggplot2, 
# and avoid scientific notation:
df$volume_mils <- df$volume/1000000

# Plotting DJIA data:
p1 <- ggplot(data = df, aes(x = date, y = close)) +  
    geom_line() +
    theme_bw() +
    labs(title = "Dow Jones Index, 2018",
         subtitle = "Market closing price",
         x = "Date", y = "Closing Price",
         caption = "Source: Yahoo! Finance") +
    geom_hline(yintercept = df$close[1], color = "blue") +
    scale_x_date(date_labels = "%b")

p2 <- ggplot(data = df, aes(x = date, y = daily_diff)) + 
    geom_line() +
    theme_bw() +
    geom_hline(aes(yintercept = 0), color = "blue") +
    labs(title = "Dow Jones Index, 2018",
         subtitle = "Daily change in market closing price",
         x = "Date", y = "Difference in Closing Price", 
         caption = "Source: Yahoo! Finance") +
    scale_x_date(date_labels = "%b")

p3 <- ggplot(data = df, aes(x = date, y = pct_change)) + 
    geom_line() +
    theme_bw() +
    geom_hline(aes(yintercept = 1.0), color = "blue") +
    labs(title = "Dow Jones Index, 2018",
         subtitle = "Percent price change; Indexed, 2 Jan 2018 = 1.00",
         x = "Date", y = "Percent", 
         caption = "Source: Yahoo! Finance") +
    scale_x_date(date_labels = "%b")

p4 <- ggplot(data = df, aes(x = date, y = level_change)) + 
    geom_line() +
    theme_bw() +
    geom_hline(aes(yintercept = 0), color = "blue") +
    labs(title = "Dow Jones Index, 2018",
         subtitle = "Actual price change; Indexed, 2 Jan 2018 = 0",
         x = "Date", y = "Amount", 
         caption = "Source: Yahoo! Finance") +
    scale_x_date(date_labels = "%b")

# Histogram:
mean_daily_diff <- mean(df$daily_diff)
sd1_daily_diff <- sd(df$daily_diff) * 1
sd2_daily_diff <- sd(df$daily_diff) * 2
sd3_daily_diff <- sd(df$daily_diff) * 3

p5 <- ggplot(data = df, aes(x = daily_diff)) + 
    geom_histogram(binwidth = 25, color = "black") + 
    geom_vline(xintercept = mean_daily_diff, color = "blue") +
    geom_vline(xintercept = sd1_daily_diff, color = "blue", linetype = 2) +
    geom_vline(xintercept = sd2_daily_diff, color = "blue", linetype = 2) +
    geom_vline(xintercept = sd3_daily_diff, color = "blue", linetype = 2) +
    geom_vline(xintercept = sd1_daily_diff * -1, color = "blue", linetype = 2) +
    geom_vline(xintercept = sd2_daily_diff * -1, color = "blue", linetype = 2) +
    geom_vline(xintercept = sd3_daily_diff * -1, color = "blue", linetype = 2) 

# Plotting daily volumes:
p6 <- ggplot(data = df, aes(x = date, y = volume_mils)) +
                 geom_col(color = "black") +
                 scale_x_discrete(labels = df$date) +
    labs(title = "Dow Jones Index, 2018",
         subtitle = "Daily Volumes for 2018; horizontal line is annual average volume",
         x = "Date", y = "Trading Volume, in millions", 
         caption = "Source: Yahoo! Finance") +
    geom_hline(yintercept = mean(df$volume_mils), color = "blue") +
    scale_x_date(date_labels = "%b")

```

What better than to start off the New Year with a blog?

This is a follow up to my previous blog about the [Dow Jones Industrial Average (DJIA)](https://www.justadatageek.com/2018/10/31/dow-s-trick-or-treat/).  The DJIA gave many people fits during the last quarter of 2018.  Here is a look at its overall performance for the year:

```{r, echo = FALSE}
grid.arrange(p1, p6, nrow = 2)
```

Another view of the DJIA's performance, this time looking at percent change and "level" change, respectively.  These charts show values that are indexed to the first trading day of the year, 2 January 2018:

```{r, echo = FALSE}
grid.arrange(p3, p4, nrow = 2)
```

I think you can agree with me when I say that the last three months were rather interesting.

I also made a plot of the distribution of the changes in the DJIA, following up on another [previous post](https://www.justadatageek.com/2018/10/11/distribution/):

```{r, echo = FALSE}
p5
```

The solid blue vertical line is the mean and the verticle broken lines are the first through third standard deviations. If you are curious, yes, the daily difference is skewed negative:

```{r, echo = FALSE}
skewness(df$daily_diff)
```

I am curious to see if the volatility of the last quarter of 2018 continues on into 2019.

On another note, I have been working on transferring several of my old Blogger posts over to this GitHub-Netlify hosted blog.  At some point, these old posts will populate into what is currently available.

I am going to leave off here.  Thank you for reading.  Good night!